Approximation as Differentiation

### 2025-06-09 main/first-order.tex
LatGal as a CMon-category.

\noindent Since right adjoints preserve limits and left adjoints preserve colimits, for any $f \adj g: X \to
Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
the join-semilattice structure $(X, \join, \bot)$.

$\LatGal$ is enriched in $\CMon$ (in fact in $\SemiLat$, the category of semilattices, but again $\CMon$ will
do for our purposes). The meet-preserving maps from $X$ to $Y$ can be added pointwise using $\meet$, since $(f
\meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto \top_Y$ as a zero;
dually the join-preserving maps from $Y$ to $X$ can be added pointwise using $\join$, with the constant map
$\bot = y \mapsto \bot_X$ as a zero. Since these constructions come in adjoint pairs, the hom-set
$\LatGal(X,Y)$ forms a commutative monoid with $0$ given by the Galois connection $\bot \adj \top$ and
addition $(f \adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$.

The projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the
product of lattices, have both upper and lower adjoints; this means that $X \times Y$, which we shall
hereafter write as $X \biprod Y$, is both a product and a coproduct, with projections $\biproj_X$ and
$\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:

>>>>

\noindent Since right adjoints preserve limits and left adjoints preserve colimits, for any $f \adj g: X \to
Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
the join-semilattice structure $(X, \join, \bot)$. The meet-preserving maps have their $\CMon$ structure
structure (in fact $\SemiLat$ structure, but again $\CMon$ is sufficient): addition of maps is pointwise via
$\meet$, since $(f \meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto
\top_Y$ as a zero. Dually  the join-preserving maps can be added pointwise using $\join$, with the constant
map $\bot = y \mapsto \bot_X$ as a zero. Since these constructions come in adjoint pairs, $\LatGal$ is
enriched in $\CMon$. The hom-set $\LatGal(X,Y)$ provides (component-wise) addition of Galois connections $(f
\adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$; the Galois connection $\bot \adj \top$ is zero
morphism, and composition is bilinear.

The projections $\pi_1$ and $\pi_2$ from the product lattice $X \times Y$, have both upper and lower adjoints;
this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$, is both a product and a
coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
